---
title: "Final Project Draft - STA 199"
subtitle: "due Thursday, Oct 29 at 11:59p"
author: "Ten Out of Ten: Arushi Bhatia, Luke Vermeer, Kevin Wang, Lauren May"
output: pdf_document
---

```{r installingPackages}
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)

install.packages("viridis")
install.packages("stopwords")
install.packages("tidyverse")
install.packages("tidytext")
install.packages("wordcloud")
install.packages("sentimentr")
install.packages("lubridate")
library(viridis)
library(stopwords)
library(tidyverse)
library(tidytext)
library(wordcloud)
library(sentimentr)
library(lubridate)
```

### Introduction
Relative to other forms of media, social media plays a much greater role in how 
people consume news in today’s technology-driven society. A study conducted in 
2016 by Pew Research points out how 62% of people get news on social media 
(Gottfried & Shearer, 2016). As a result, social media plays an integral role 
in politics, as the presence of a specific subset of information on a user’s 
feed can influence the way they categorize and see the world around them. With 
more and more individuals on social media such as Twitter, the role that this 
platform can play is significantly greater than before. 

Social media, and Twitter in particular, have become an increasingly large 
part of the political landscape in the wake of Donald Trump's 2016 election. 
Trump has been active on Twitter prior to and during his presidency and uses 
the platform as a tool to communicate with his constituency in real time, 
posting updates about policy, campaigning, and his feelings on everything 
from members of Congress to celebrities. He is one of the first politicians 
to use social media this frequently and has personally referred to his use 
of Twitter as “modern day presidential” (Trump, 2017). 

Donald Trump’s Twitter also has “unpresidented” reach among the American public, 
boasting a follower base of over 87 million. This makes him the second
most-followed political personality and sixth most-followed overall account on 
Twitter (Wikipedia, 2020). On top of this, Trump’s Twitter also receives 
significant attention in the media. Over 850,000 news articles have referenced 
his Twitter use since 2016 and 31% of his Tweets since then have received 
individual media coverage (Real Clear Politics, 2019).

Because Trump uses Twitter to convey his political agendas in short blurbs, 
analyzing his Tweets can give a unique insight into the way that he thinks. 
The research question we will be exploring is how President Trump’s sentiment 
in his Tweets and its reception by his large Twitter following (popularity) 
varies across people, organizations, and policies. Our hypothesis regarding 
our research question is that Trump’s tweets have negative sentiments towards 
opposing groups and policies and positive sentiments towards groups and policies 
that align with his beliefs. In addition, we believe the tweets with negative 
sentiments will be slightly more popular in terms of favorites and retweets.

### Data description
The dataset was extracted from a website - TrumpTwitterArchive.com. The original 
curator of the data created their own Twitter scraper in order to obtain the 
data. They utilized Python, Selenium (which is a software suite that allows the 
automation of tests utilizing web browsers), and Tweepy (a Python library for 
accessing the Twitter API). Since Twitter makes it challenging to scrape all of 
a user’s Tweets in one go, the way to get around this is to individually search 
for a specific day and extract all the Tweets from that user on that specific 
day. To do this manually would take ages, but the scraper that the curator built 
allows for automated accessing for any desired day and also a range of days. 
The scraper then obtains the Tweet ID, which contains all of the metadata of the 
Tweet, and then uses the metadata to obtain all the other information about the 
Tweet (such as the text, timestamp, number of favorites, etc.). This other 
information is then compiled into a dataset, which is made available to the 
public. This dataset is updated every minute, which also means that deleted 
Tweets would most likely also appear in this dataset. 
 
This data set entitled trumpTweets includes 53,697 observations. Each individual 
observation is one of President Donald Trump’s tweets. The original dataset 
contains 7 variables: source, text, created_at, retweet_count, favorite_count, 
is_retweeted, id_str. The descriptions of each of the original variables is 
given below. 

- source: Original source where tweet was posted 
- text: text of the tweet 
- created_at: Date and time the tweet was posted/created, provides context
- retweet_count: number of retweets 
- favorite_count: number of favorites 
- is_retweeted: whether or not the tweet was originally posted on a different 
account and Trump retweeted 
- id_str: The scrape.py script collects tweet ids. If you know a tweet's id 
number, you can get all the information available about that tweet using Tweepy 
- text, timestamp, number of retweets / replies / favorites, geolocation, etc. 

### Methodology 

Our first step in beginning to analyze the relationships between subject, 
sentiment, and reception of President Donald Trump’s tweets was to manipulate 
the data set, creating some new variables that we could use for analysis. We 
started by creating a set of identifier variables for different people that his 
tweets might be about. These variables were created using a mutate command to 
set them as either 1 or 0, 1 if the person was mentioned in a tweet, and 0 if 
they weren’t. We created these variables for a total of 11 relevant political 
figures: Barack Obama, Joe Biden, Hillary Clinton, Alexandria Ocasio-Cortez, 
Nancy Pelosi, Kamala Harris, Mike Pence, Mitch McConnell, Amy Coney Barrett, 
and Nikki Haley. For each of the person identifier variables, we searched for 
people’s full names, their Twitter handles, and commonly-used nicknames to 
determine whether or not a particular tweet was about them. We also repeated the 
same process to create topical identifier variables for a range of subjects that 
are prevalent in the nation’s political discourse. The topics were as follows: 
COVID-19, climate change, abortion, the Black Lives Matter movement, guns, news, 
immigration, Russia, and the United States. For each topic we searched for the 
name of the thing outright (such as “BLM” or “climate change”), as well as words 
and phrases that are commonly used in conjunction with these topics. For 
example, tweets that mentioned ICE or the term “border wall” were categorized 
under immigration, and tweets about CNN, FOX News and “media” all went under the 
“news” topic. Next, using the pivot command in r, we created a variable for 
“person,”  a categorical variable that told us which of the identifier variables 
for people were equal to “1” for each tweet. The output was a variable that had 
a value of either the name of one of the 11 people we searched for or “other.” 
From the new “person” variable, we created two new variables called party and 
gender. The party variable separated the 11 people we were looking at into 
either democrats or republicans and assigned tweets about them to the 
appropriate party category. We did the same for the gender of each of the 
people giving tweets about these figures a value of either “male” or “female.” 
Next, we used the package sentimentr to identify the sentiment of each of 
Trump’s tweets. The package examines each tweet, identifying positive and 
negative words based on a lexicon made by the package’s creator Tyler Rinker. 
It then aggregates a sentiment score for the tweet, assigning a value of -1 to 
words it deems negative, 1 for positive words and 0 for neutral words. This 
results in an overall sentiment score. For each tweet, we created a new variable 
called ave_sentiment that contained the raw, numeric sentiment score of the 
text.  We then used the mutate command to create a new variable called posNeg 
that grouped sentiment scores into three categories: positive, neutral, or 
negative. Finally, we used the separate command in r to break up the variable 
“created_at” into two separate variables for date and time. After this, each 
observation had a date variable in the mm/dd/yy format and a time variable in 
the 24-hour time format. After these manipulations, the data set contained the 
following variables:

- source: the version of Twitter used to share the tweet. For example “Twitter 
for iPhone.”
- text: the full text of the tweet.
- created_at: gives the date (in mm/dd/yy format) and time (in 24-hour format) 
at which the tweet was published.
- retweet_count: number of retweets that the tweet received.
- favorite_count: number of favorites that the tweet received.
- is_retweet: assigns a value of either true or false for whether the tweet is 
originally written by Trump or is a retweet of someone else.
- id_str: The scrape.py script collects tweet ids. If you know a tweet's id 
you can get all the information available about that tweet using Tweepy - text, 
timestamp, number of retweets / replies / favorites, geolocation, etc. 
- obama: 1 or 0 for whether the tweet talks about Barack Obama.
- biden: 1 or 0 for whether the tweet talks about Joe Biden.
- pelosi: 1 or 0 for whether the tweet talks about Nancy Pelosi.
- kamala: 1 or 0 for whether the tweet talks about Kamala Harris.
- hillary: 1 or 0 for whether the tweet talks about Hillary Clinton.
-aoc: 1 or 0 for whether the tweet talks about Alexandria Ocasio-Cortez.
- pence: 1 or 0 for whether the tweet talks about Mike Pence.
- mcconnell: 1 or 0 for whether the tweet talks about Mitch McConnell.
- fauci: 1 or 0 for whether the tweet talks about Dr.  Anthony Fauci.
- amy: 1 or 0 for whether the tweet talks about Amy Coney Barrett.
- nikki: 1 or 0 for whether the tweet talks about Nikki Haley.
- covid: 1 or 0 for whether the tweet talks about COVID-19.
- climateChange: 1 or 0 for whether the tweet talks about climate change.
- abortion: 1 or 0 for whether the tweet talks about abortion.
- blm: 1 or 0 for whether the tweet talks about the Black Lives Matter movement.
- guns: 1 or 0 for whether the tweet talks about guns.
- news: 1 or 0 for whether the tweet talks about news media.
- usa: 1 or 0 for whether the tweet talks about the United States.
- russia: 1 or 0 for whether the tweet talks about Russia.
- immigration: 1 or 0 for whether the tweet talks about immigration.
- person: categorical variable for the name of the person that the tweet is 
about (from among the 11 looked at in this analysis).
- party: political party of the person who is talked about in the tweet 
(either democrat or republican).
- gender: gender of the person who is talked about in the tweet (either male or 
female).
- ave_sentiment: numeric sentiment score of the tweet.
- posNeg: categorical sentiment of the tweet (either positive, neutral or 
negative).
- ate: date that the tweet was posted (mm/dd/yy).
- time: time that the tweet was posted (24-hour format).

### Glimpse of data
```{r packages-data, warning=FALSE, message=FALSE}
trumpTweets <- read_csv("data/trumpTweetCSV.csv") 
glimpse(trumpTweets)
```
```{r people}
trumpTweets <- trumpTweets %>%  
   mutate(obama = case_when((grepl("Obama", text , ignore.case = TRUE) | 
                              grepl("Barack", text , ignore.case = TRUE)) & 
                            !(grepl("Michelle", text , ignore.case = TRUE))~ 1,
                            TRUE ~0))

trumpTweets <- trumpTweets %>%  
   mutate(biden = case_when((grepl("Biden", text , ignore.case = TRUE) | 
                              grepl("Joe", text , ignore.case = TRUE)) ~ 1,
                            TRUE ~0))

trumpTweets <- trumpTweets %>%  
   mutate(pelosi = case_when((grepl("Pelosi", text , ignore.case = TRUE) | 
                              grepl("Nancy", text , ignore.case = TRUE)) &
                            ! grepl("@NancyMace", text, ignore.case = TRUE)~ 1,
                            TRUE ~0))


trumpTweets <- trumpTweets %>%  
   mutate(kamala = case_when((grepl("Kamala", text , ignore.case = TRUE))~ 1,
                            TRUE ~0))


trumpTweets <- trumpTweets %>%  
   mutate(hillary = case_when((grepl("Hillary", text , ignore.case = TRUE) | 
                              grepl("Clinton", text , ignore.case = TRUE)) & 
                        !(grepl("Bill Clinton", text , ignore.case = TRUE))~ 1,
                            TRUE ~0))

trumpTweets <- trumpTweets %>%  
   mutate(aoc = case_when((grepl("AOC", text , ignore.case = TRUE) | 
                              grepl("Ocasio", text , ignore.case = TRUE) |
                            grepl("Cortez", text , ignore.case = TRUE)|
                             grepl("Alexandria", text , ignore.case = TRUE))~ 1,
                            TRUE ~0))

trumpTweets <- trumpTweets %>%  
   mutate(pence = case_when((grepl("Pence", text , ignore.case = TRUE)) ~ 1,
                            TRUE ~0))

trumpTweets <- trumpTweets %>%  
   mutate(nikki = case_when((grepl("Nikki Haley", text , ignore.case = TRUE)| 
                              grepl("Nikki", text , ignore.case = TRUE)) ~ 1,
                            TRUE ~0))


trumpTweets <- trumpTweets %>%  
  mutate(mcconnell = case_when(((grepl("McConnell", text , ignore.case = TRUE) | 
                              grepl("Mitch", text , ignore.case = TRUE) |
                            grepl("@Team_Mitch", text , ignore.case = TRUE)) &
                        ! grepl("@mitchellvii", text , ignore.case = TRUE))~ 1,
                            TRUE ~0))

trumpTweets <- trumpTweets %>%  
   mutate(fauci = case_when((grepl("Fauci", text , ignore.case = TRUE)) ~ 1,
                            TRUE ~0))

trumpTweets <- trumpTweets %>%  
   mutate(amy = case_when((grepl("Coney Barrett", text , ignore.case = TRUE) | 
                             grepl("ACB", text , ignore.case = TRUE)| 
                             grepl("Judge Amy", text , ignore.case = TRUE)) ~ 1,
                            TRUE ~0))
```


```{r topics}
trumpTweets <- trumpTweets %>%  
   mutate(covid = case_when((grepl("COVID", text , ignore.case = TRUE) | 
                              grepl("Corona", text , ignore.case = TRUE) |
                      grepl("Chinese virus", text , ignore.case = TRUE) |
                        grepl("China virus", text , ignore.case = TRUE) |
                        grepl("Chinese plague", text, ignore.case = TRUE) |
                        grepl("Wuhan virus", text , ignore.case = TRUE) |
                        grepl("Kung flu", text , ignore.case = TRUE)|
                        grepl("cdc", text , ignore.case = TRUE)|
                        grepl("fauci", text , ignore.case = TRUE)) ~ 1,
                            TRUE ~0))

trumpTweets <- trumpTweets %>%  
   mutate(climateChange = case_when((grepl("Climate change", text , 
                                           ignore.case = TRUE) | 
                          grepl("Global warming", text , ignore.case = TRUE) |
                      grepl("Environment", text , ignore.case = TRUE) |
                grepl("Paris Accord", text , ignore.case = TRUE) |
                 grepl("Paris Climate Agreement", text, ignore.case = TRUE) |
                      grepl("Pollute", text , ignore.case = TRUE) |
                      grepl("Pollution", text , ignore.case = TRUE)|
                      grepl("Greta Thunberg", text, ignore.case = TRUE)|
                      grepl("Greta", text, ignore.case = TRUE)) ~ 1,
                            TRUE ~0))

trumpTweets <- trumpTweets %>%  
   mutate(abortion = case_when((grepl("Abortion", text , ignore.case = TRUE) | 
                        grepl("Planned parenthood", text , ignore.case = TRUE) |
                      grepl("Abort", text , ignore.case = TRUE) |
                        grepl("pro-life", text , ignore.case = TRUE) |
                        grepl("pro-choice", text , ignore.case = TRUE)) ~ 1,
                            TRUE ~0))


trumpTweets <- trumpTweets %>%  
   mutate(blm = case_when((grepl("BLM", text , ignore.case = TRUE) | 
                       grepl("Black lives matter", text , ignore.case = TRUE) |
                      grepl("loot", text , ignore.case = TRUE) |
                        grepl("riot", text , ignore.case = TRUE)) ~ 1,
                            TRUE ~0))

trumpTweets <- trumpTweets %>%  
   mutate(guns = case_when((grepl("NRA", text , ignore.case = TRUE) | 
                          grepl("Second amendment", text , ignore.case = TRUE) |
                      grepl("gun", text , ignore.case = TRUE) |
                        grepl("bear arms", text , ignore.case = TRUE) |
                        grepl("rifle", text , ignore.case = TRUE) ) ~ 1,
                            TRUE ~0))

trumpTweets <- trumpTweets %>%  
   mutate(news = case_when((grepl("news", text , ignore.case = TRUE) |
                      grepl("CNN", text , ignore.case = TRUE) |
                        grepl("Fox", text , ignore.case = TRUE) |
                        grepl("media", text , ignore.case = TRUE)) ~ 1,
                            TRUE ~0))

trumpTweets <- trumpTweets %>%  
   mutate(usa = case_when((grepl("usa", text , ignore.case = TRUE) | 
                              grepl("america", text , ignore.case = TRUE) |
                      grepl("united states", text , ignore.case = TRUE) |
                        grepl("our country", text , ignore.case = TRUE)) ~ 1,
                            TRUE ~0))

trumpTweets <- trumpTweets %>%  
   mutate(russia = case_when((grepl("russia", text , ignore.case = TRUE) | 
                              grepl("putin", text , ignore.case = TRUE)| 
                            grepl("communism", text , ignore.case = TRUE)) ~ 1,
                            TRUE ~0))

trumpTweets <- trumpTweets %>%  
   mutate(immigration = case_when((grepl("immigrants", text , 
                                         ignore.case = TRUE) | 
                              grepl("immigration", text , ignore.case = TRUE) |
                      grepl("foreigner", text , ignore.case = TRUE) |
                        grepl("open border", text , ignore.case = TRUE) |
                        grepl("the wall", text , ignore.case = TRUE) |
                        grepl("border wall", text , ignore.case = TRUE)|
                        grepl("Mexicans", text , ignore.case = TRUE)|
                        grepl("refugees", text , ignore.case = TRUE)|
                        grepl("ice", text , ignore.case = TRUE)|
                        grepl("border patrol", text , ignore.case = TRUE)|
                        grepl("cages", text , ignore.case = TRUE)) ~ 1,
                            TRUE ~0))


trumpTweets$element_id <- 1:nrow(trumpTweets) 
```


```{r creatingSentimentDataset, message=FALSE}
trumpTweetsSentiment <- sentiment_by(trumpTweets$text)
tweetsWithSentiment <- left_join(trumpTweets, trumpTweetsSentiment, 
                                 by = "element_id")

tweetsWithSentiment <- tweetsWithSentiment %>% 
   mutate(posNeg = case_when(ave_sentiment>0 ~ "positive", 
                             ave_sentiment<0 ~ "negative", 
                             ave_sentiment ==0 ~ "neutral"))

tweetsWithSentimentWithoutLink <- tweetsWithSentiment %>% 
  filter(!grepl("^https*", text))
```
```{r wordCloud, message=FALSE, echo=FALSE}
justWords <- tweetsWithSentimentWithoutLink %>%
  unnest_tokens(word, text)

freqWords <- justWords %>%
  anti_join(get_stopwords(source = "smart"))  %>%
  count(word) %>%
  arrange(desc(n)) %>% 
   filter(!(word=="t.co" | word=="realdonaldtrump"| word=="https" | word=="rt"| 
      word=="amp" | word=="http" | word =="10" | word=="2015" | word=="2016"))
wordCloud <- freqWords %>% with(wordcloud(word, n, max.words = 100))
```

Figure 1: Word cloud of the top 100 words Donald Trump uses in his Tweets. Some
of the words that were not part of the actual text of the Tweet or were
artifacts of Tweets (such as URLs, account names, and years) were removed to 
clean the data.

```{r wordFreqGraph, message=FALSE, echo=FALSE}
freqChart <- freqWords %>%
  top_n(20) %>%
  ggplot(aes(fct_reorder(word, n), n)) +
    geom_col() +
    coord_flip() + 
    theme_minimal() +
    labs(title = "Frequency of words in Donald Trump's Tweets",
         subtitle = "Top 20 word displayed",
         y = "Frequency",
         x = "Words")

freqChart
```

Figure 2: Frequency of Donald Trump's top 20 words that are used in his Tweets.
Just like in Figure 1, some of the words that were not part of the actual text 
of the Tweet or were artifacts of Tweets (such as URLs, account names, and 
years) were removed to clean the data.

```{r creatingNewDataFrameWithPeople}

peopleData <- tweetsWithSentimentWithoutLink %>%
  pivot_longer(obama:amy, names_to = "person", values_to = "existenceOfPerson")

peopleData <- peopleData %>% filter(existenceOfPerson==1)

ggplot(data=peopleData, aes(x=person, fill = factor(posNeg))) +
  geom_bar(position = "fill") + 
  labs(title="Lowest proportion of Tweets with negative sentiment were Tweets
  mentioning Amy Coney Barrett and Mike Pence",
                  x="Politician",y ="Percentage of Tweets", fill="Sentiment") + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
Figure 3: Illustrates the proportion of Tweets with average negative (<0), 
neutral (=0), and positive (>0) sentiment that mention a specific politician.

```{r partyAndGender}
peopleData <- peopleData %>% 
   mutate(party=case_when((person == "obama" |
               person == "biden"|
               person == "kamala" | 
               person == "pelosi" | 
               person == "aoc"|
                  person == "hillary") ~ "Democratic",
               (person == "pence" |
                   person == "amy" |
                   person == "niki" |
                  person == "mcconnell") ~ "Republican"))

peopleData <- peopleData %>% 
   mutate(gender=case_when((person=="obama" |
              person=="biden"|
               person=="pence"|
               person=="fauci"| 
               person=="mcconnell") ~ "Male",
               (person=="kamala" |
                   person=="pelosi" |
                   person=="aoc"|
                   person=="amy" |
                   person=="nikki"|
                  person=="hillary") ~ "Female"))

ggplot(data = peopleData %>% 
          filter(party=="Democratic" | party == "Republican"), 
       mapping = aes(x = party, y = ave_sentiment,color=gender)) +
   geom_boxplot() +
labs(title = "Overall more positive sentiment in Tweets mentioning Republicans",
       subtitle="Generally lower sentiment for Democratic females vs. males, 
   generally higher sentiment for Republican females vs. males",
        x = "Party", y = "Tweet Average Sentiment Score") + 
   geom_hline(yintercept=0, linetype="dashed", color = "red")
```

Figure 4: Boxplot showing the average sentiment scores for Tweets, divided by
political party of the person mentioned in the Tweet, and split within
party to show any differences in how he talks about politicians of
different genders within either the Democratic or Republican party.


```{r hypTest}
dems <- peopleData %>%
  filter(party=="Democratic")

repubs <- peopleData %>%
  filter(party=="Republican")

t.test(dems$ave_sentiment, 
       repubs$ave_sentiment,
       mu = 0,
       var.equal = FALSE,
       alternative = "less",
       conf.level = 0.95)
```



```{r topicsAcrossTime}
tweetsWithSentimentDateTime <- tweetsWithSentimentWithoutLink %>%
 separate(created_at, c("date", "time"), sep = " ")

tweetsWithSentimentDateTimeFormat <- tweetsWithSentimentDateTime %>%
 mutate(date = as.Date(date, tryFormats = c("%m/%d/%y")))%>%
 pivot_longer(covid:immigration, names_to = "topic", values_to = 
                 "existenceOfTopic") %>%
 filter(existenceOfTopic == 1)

tweetsWithSentimentDateTimeFormatGroup <- tweetsWithSentimentDateTimeFormat %>%
   mutate(month = format(date, "%m"), year = format(date, "%Y")) %>%
   group_by(topic, year) %>% 
   summarize(avgSentimentForTopic = mean(ave_sentiment))

temp <- tweetsWithSentimentDateTimeFormatGroup %>% filter(topic=="climateChange"
   | topic=="news" | topic == "guns" | topic == "immigration") %>% 
   filter(!is.na(year))

ggplot(temp, 
       aes(x= year, y = avgSentimentForTopic, color = topic, 
           group=topic)) + theme(plot.margin = unit(c(1,0,1,0), "cm")) + 
   geom_point() + 
   geom_line() + 
   geom_vline(xintercept = "2016", linetype="dotted", 
                color = "blue", size=1.5) +
   labs(title = "Average sentiment for specific topics across the years",
       subtitle="Blue line indicates Trump's election, 
     immigration and news follow similar patterns",
   x = "Year", y = "Tweet Average Sentiment Score Per Topic", color="Topic") + 
   geom_hline(yintercept=0, linetype="dashed", color = "red")
```

Figure 5: Shows change of average sentiment of Tweets referring to a specific 
topic across the years. Only showing a few topics, we selected the ones
that have been Tweeted about for a long time. The vertical blue line indicates
when Trump first became elected.

```{r hypTest2}
yearVar <- tweetsWithSentimentDateTimeFormat  %>%
   mutate(year = year(date))

#CLIMATE CHANGE
pre2016CC <- yearVar %>%
  filter(year==2016 & topic=="climateChange")

post2016CC <- yearVar %>%
  filter(year>=2016 & topic=="climateChange")

t.test(pre2016CC$ave_sentiment, 
       post2016CC$ave_sentiment,
       mu = 0,
       var.equal = FALSE,
       alternative = "two.sided",
       conf.level = 0.95)

#GUNS 
pre2016guns <- yearVar %>%
  filter(year==2016 & topic=="guns")

post2016guns <- yearVar %>%
  filter(year>=2016 & topic=="guns")

t.test(pre2016guns$ave_sentiment, 
       post2016guns$ave_sentiment,
       mu = 0,
       var.equal = FALSE,
       alternative = "two.sided",
       conf.level = 0.95)


#IMMIGRATION 
pre2016imm <- yearVar %>%
  filter(year==2016 & topic=="immigration")

post2016imm <- yearVar %>%
  filter(year>=2016 & topic=="immigration")

t.test(pre2016imm$ave_sentiment, 
       post2016imm$ave_sentiment,
       mu = 0,
       var.equal = FALSE,
       alternative = "two.sided",
       conf.level = 0.95)


#NEWS
pre2016News <- yearVar %>%
  filter(year==2016 & topic=="news")

post2016News <- yearVar %>%
  filter(year>=2016 & topic=="news")

t.test(pre2016News$ave_sentiment, 
       post2016News$ave_sentiment,
       mu = 0,
       var.equal = FALSE,
       alternative = "two.sided",
       conf.level = 0.95)
```

```{r chiSquare}
polarization_tweets <- tweetsWithSentiment %>%
   mutate(polarization = abs(ave_sentiment)) %>%
   mutate(retweet_size = case_when(retweet_count <= 20000 ~ 1,
      retweet_count > 20000 & retweet_count <= 40000 ~ 2,
      retweet_count > 40000 & retweet_count <= 60000 ~ 3,
      retweet_count > 60000 & retweet_count <= 80000 ~ 4,
      retweet_count > 80000 & retweet_count <= 100000 ~ 5, 
      retweet_count > 100000 ~ 6))

chisq.test(table(polarization_tweets$posNeg, 
   polarization_tweets$retweet_size))

```
```{r linModel}

tweetsWithSentimentWithoutLink <- tweetsWithSentimentWithoutLink%>%
   filter(retweet_count!=0 & !is.na(retweet_count)) %>%
   mutate(log_retweet_count = log(retweet_count))

person_retweets <- lm(log_retweet_count ~ obama + biden + pelosi + kamala + hillary 
   + aoc + pence + nikki + mcconnell + amy + fauci, 
   data = tweetsWithSentimentWithoutLink)

tidy(person_retweets)


aug <- augment(person_retweets)

ggplot(data=aug, aes(x=.resid)) + geom_histogram()

ggplot(aug, mapping = aes(sample = .resid)) + 
   stat_qq() + stat_qq_line()

ggplot(data=aug, aes(x=.fitted, y=.resid)) + geom_point()+ 
   geom_hline(yintercept=0, col="red")

#fix this, correlation between people?
trees %>% summarize(correlation = cor(Height, Girth))




topic_retweets <- lm(log_retweet_count ~ covid + climateChange + abortion + blm + 
   guns + news + usa + russia + immigration, 
   data = tweetsWithSentimentWithoutLink)

tidy(topic_retweets)

aug <- augment(topic_retweets)

ggplot(data=aug, aes(x=.resid)) + geom_histogram()

ggplot(aug, mapping = aes(sample = .resid)) + 
   stat_qq() + stat_qq_line()

ggplot(data=aug, aes(x=.fitted, y=.resid)) + geom_point()+ geom_hline(yintercept=0, col="red")

#fix this, correlation between topics?
trees %>% summarize(correlation = cor(Height, Girth))

```


```{r linModel2}
person_retweets <- lm(retweet_count ~ obama + biden + pelosi + kamala + hillary 
                      + aoc + pence + nikki + mcconnell + amy + fauci, 
                 data = tweetsWithSentimentWithoutLink)

tidy(person_retweets)


aug <- augment(person_retweets)

ggplot(data=aug, aes(x=.resid)) + geom_histogram()

ggplot(aug, mapping = aes(sample = .resid)) + 
  stat_qq() + stat_qq_line()

ggplot(data=aug, aes(x=.fitted, y=.resid)) + geom_point()+ 
   geom_hline(yintercept=0, col="red")

#fix this, correlation between people?
trees %>% summarize(correlation = cor(Height, Girth))




topic_retweets <- lm(retweet_count ~ covid + climateChange + abortion + blm + 
                        guns + news + usa + russia + immigration, 
                 data = tweetsWithSentimentWithoutLink)

tidy(topic_retweets)

aug <- augment(topic_retweets)

ggplot(data=aug, aes(x=.resid)) + geom_histogram()

ggplot(aug, mapping = aes(sample = .resid)) + 
  stat_qq() + stat_qq_line()

ggplot(data=aug, aes(x=.fitted, y=.resid)) + geom_point() + 
   geom_hline(yintercept=0, col="red")

#fix this, correlation between topics?
trees %>% summarize(correlation = cor(Height, Girth))

```
### Results
(Figure 1)
The purpose of this visualization is to analyze what words are most commonly 
found in Donald Trump’s tweet. Using a word cloud, the most common words are 
displayed with the most frequent words being displayed larger. Given the 
results, we can see that some of the words he most commonly uses are Trump, 
president, people, country, america, and time. While these words don’t 
necessarily indicate the full scope of what he discusses on his Twitter 
account, we are able to draw the conclusion that most of his tweets are in 
regard to his role as the President of the United States. We believed it would 
be interesting to see if a common theme would develop among his most common 
words, which is partially indicated in our results. 

(Figure 3)
Figure 3 shows the relative proportions of tweets with positive, neutral and 
negative overall sentiment that Donald Trump has shared about each of the people 
covered in our dataset. By creating a bar plot, we were able to actively compare 
the sentiments Trump tends to have towards specific people making it easier to 
visualize the results. Overall, it appears that Alexadria Ocasio-Cortez, Hillary 
Clinton, and Nancy Pelosi recieved the highest proportion of tweets with 
negative sentiment. Amy Coney Barrett had the lowest proportion of negative 
tweets, followed by Vice President Mike Pence. On the opposite side of the 
spectrum, Senate Majority Leader Mitch McConnell had the highest proportion of 
positive tweets while Amy Coney Barrett, Mike Pence, and Nikki Haley also 
received a significant majority of tweets with a positive sentiment. These 
results were particularly interesting because it shows that Trump tends to 
tweet about people of the same party with positive sentiment more frequently 
compared to the people of the democratic party. 

(Chi Square Test)
The variable, polarization, is defined as how intense the sentiment of 
Donald Trump’s tweet is. This hypothesis test will analyze if the polarization 
of the tweet and the amount of retweets are dependent or independent of each 
other. We chose to analyze this because we believed it to indicate the public’s 
response to his tweets. We can draw the conclusion that the public reaction is 
dependent on how intense of a sentiment Trump expresses if we reject the null 
hypothesis.

We are performing a hypothesis test using a chi-square test. We are performing 
this hypothesis test at the 0.05 significance levels. 
Here are our null and alternative hypotheses: 

$H_0$: The polarization of the tweet and the amount of retweets are independent; 
there is no association between the two variables. 

$H_1$: The polarization of the tweet  and the amount of retweets are NOT 
independent; there is an association between the two variables 

Under the assumption that the null hypothesis is true, the chi-square test 
statistic follows a Chi-square distribution with degrees of freedom equal to 
39554.

The value of our test statistic, chi-square, is 41580. 

The p-value of our chi-square test was 6.949e-13, which is less than the 
significance level of 0.05. This means we reject the null hypothesis. We are 
concluding that there is sufficient evidence to suggest that the presence of 
throat pain and ASA classification are NOT independent or that there is 
sufficient evidence to suggest that there is an association between the two 
variables. More specifically, this hypothesis test shows that the amount of 
retweets are dependent on how polarizing President Trump’s tweets are. 


The purpose of these visualizations is to indicate the different sentiment he 
displays towards male versus female politicians. We decided to divide this into 
two visualizations by party affiliation because we believed the results could be 
skewed by differing political beliefs, and since we just wanted to learn if he 
had a different sentiment towards strictly males or females, we figured this was 
the best way to isolate the analysis of the gender variable. This visualization 
of the Democrats shows the average sentiment score among the Democratic 
politicians, showing specifically the difference in sentiment between male and 
female politicians using a box plot. This visualization shows that when Donald 
Trump discusses Democratic politicians, he has a similar sentiment towards male 
and female, which is fairly close to “neutral”. We also noticed that for this 
visualization, the male Democrats had a larger range of sentiments than the 
female Democrats. In the visualization of male and female Republicans, the 
females had a slightly higher average sentiment compared to their male 
counterparts. This could indicate that he talks about males in his party in a 
less positive manner as compared to the females. We also recognized that, again, 
the male Republicans had a larger range of sentiments than the female 
Republicans, showing that his sentiments are more likely to fluctuate when 
discussing a male in his tweets.

(Topics over time)
The purpose of this visualization is to see if Donald Trump’s sentiments towards 
specific “hot button topics” change over time. We compiled all the tweets into 
a line plot with time as the x-axis and average sentiment as the y-axis. By 
taking the average sentiment of each topic by year, we were able to create this 
visualization and see how his sentiments in his tweets changed when discussing 
the topics. 
### Sources
http://www.trumptwitterarchive.com/about

https://www.journalism.org/2016/05/26/news-use-across-social-media-platforms-
2016/

https://www.realclearpolitics.com/articles/2019/09/11/numbers_show_how_trumps_
tweets_drive_the_news_cycle_141217.html

https://en.wikipedia.org/wiki/List_of_most-followed_Twitter_accounts

https://twitter.com/realDonaldTrump?ref_src=twsrc%5Egoogle%7Ctwcamp%5Eserp%7
Ctwgr%5Eauthor



